---
title: "Mastery Assignment Jake Arlow"
output: html_notebook
---
```{r}
library(BIEN)
library(plyr)
library(labdsv)
library(vegan)
library(divDyn)
```

For this mastery assignment I wanted to compare modern plant diversity to Jurassic plant diversity in New Jersey. For the modern data I compiled a list using Plant-O-Matic, and for the Jurassic lists I used both PaleoDB and the article "Paleontology and paleoenvironments of early jurassic age strata in the walter kidde Dinosaur Park (New Jersey, USA)

I am going to compare the diversity indices of modern and Jurassic New Jersey.


```{r}
modern_list = read.csv("modern.csv")
jurassic_list = read.csv("jurassic.csv")
modern_species = paste(modern_list$Scientific.Name)
jurassic_species = paste(jurassic_list$Scientific.Name)

#print
modern_species
jurassic_species
```

I am using BIEN for the modern abundance data, and I found an R package called "divDyn" that is supposed to pull from the PaleoDB, but I am having trouble getting it to work. DivDyn "describes sampling and diversity dynamics of fossil occurrence datasets (e.g. from the Paleobiology Database)" and calculates taxonomic richness. 

```{r}
modern_ab = count(modern_species)
modern_ab = data.frame("Modern", modern_ab)
names(modern_ab)<-c("ListID", "SpeciesID", "Freq")
modern_ab

jurassic_ab = count(jurassic_species)
jurassic_ab = data.frame("Jurassic", app_ab)
names(jurassic_ab)<-c("ListID", "SpeciesID", "Freq")
jurassic_ab
```

Matrify

```{r}
all_ab = rbind(modern_ab, jurassic_ab)
all_matrix = matrify(all_ab)
all_matrix
```

Diversity/similarity indices:

```{r}
simpson<-diversity(all_matrix, index = "simpson")
shannon<-diversity(all_matrix, index = "shannon")
print("Simpson Index: ", quote = FALSE)
simpson
print("Shannon Index: ", quote = FALSE)
shannon
jaccard = vegdist(all_matrix, "jaccard")
print("Jaccard Index: ", quote = FALSE)
jaccard
bray = vegdist(all_matrix, "bray")
print("Bray Index: ", quote = FALSE)
bray
```
I'm not sure how well this worked, but it makes sense that the Jurassic plant list has lower diversity for a few reasons. First of all, not every species that occurred in Jurassic New Jersey has been recorded or fossilized. Also, there were no angiosperms yet in the Jurassic, only gymnosperms. 
