---
title: "Portfolio 2"
output: html_notebook
---

Portfolio 2, Part 2.3 
By Jake Arlow

```{r}
library(BIEN)
library(plyr)
library(labdsv)
library(vegan)
```

I'm getting occurence data from Oregon. The three counties are Benton (where OSU is located), Multnomah (Portland) and Malheur

```{r}
benton <- BIEN_occurrence_county(country="United States", state="Oregon", county = "Benton")
benton
multnomah <- BIEN_occurrence_county(country="United States", state="Oregon", county = "Multnomah")
multnomah
malheur <- BIEN_occurrence_county(country="United States", state="Oregon", county = "Malheur")
malheur
```

Occurence/abundance data for the counties: 

```{r}
# occurrence data
benton_occ = benton[c(1)]
multnomah_occ = multnomah[c(1)]
malheur_occ = malheur[c(1)]

# abundance data
benton_ab = count(benton_occ)
multnomah_ab = count(multnomah_occ)
malheur_ab = count(malheur_occ)

# checking data
benton_occ
benton_ab
```

Cleaning up occurence/abundance data: 

```{r}

# renaming data frame
benton_ab = data.frame("benton", benton_ab)
multnomah_ab = data.frame("multnomah", multnomah_ab)
malheur_ab = data.frame("malheur", malheur_ab)

# renaming columns
names(benton_ab)<-c("ListID", "SpeciesID", "Freq")
names(multnomah_ab)<-c("ListID", "SpeciesID", "Freq")
names(malheur_ab)<-c("ListID", "SpeciesID", "Freq")

# printing data frames
benton_ab
multnomah_ab
malheur_ab
```


```{r}
#binding data frames
counties = rbind(benton_ab, multnomah_ab, malheur_ab)
counties
```

Now we're gonna matrify (such a fun word).

```{r}
counties_matrix<-matrify(counties)
counties_matrix
```

Computing the two different diversity indices.

```{r}
simpson_index<-diversity(counties_matrix, index = "simpson")
shannon_index<-diversity(counties_matrix, index = "shannon")
print("Simpson Index: ", quote = FALSE)
simpson_index
print("Shannon Index: ", quote = FALSE)
shannon_index
```

The Simpson values for the three counties are very similar. Malheur has a higher Shannon index than Benton and Multnomah counties. The Simpson index is also slightly higher for Malheur county, which means that there is a slightly higher diversity.  

Bray and Jacard indices to look at the difference b/t counties

```{r}
bray_index = vegdist(counties_matrix, "bray")
jaccard_index = vegdist(counties_matrix, "jaccard", binary=T)
print("Bray Index: ", quote = FALSE)
bray_index
print("Jaccard Index: ", quote = FALSE)
jaccard_index
```

The higher the Bray index, the fewer shared species. Malheur county is more different from Benton county than Multnomah is. 

Looking at diversity from plant lists assembled for portfolio. The Welikia list is from 51st between 8th and 9th because that's near where my Grandma lives, a list from Plant-o-Matic (Corvallis Oregon), and then plants from my article. 

```{r}
welikia_list = read.csv("welikia.csv")
app_list = read.csv("plant-o-matic_plants.csv")
article_list = read.csv("plants_article.csv")
welikia_species = paste(welikia_list$Scientific.Name)
app_species = paste(app_list$Scientific.Name)
article_species = paste(article_list$Scientific.Name)

#print
welikia_species
app_species
article_species
```

Abundance data:

```{r}
welikia_ab = count(welikia_species)
welikia_ab = data.frame("Welikia", welikia_ab)
names(welikia_ab)<-c("ListID", "SpeciesID", "Freq")
welikia_ab

app_ab = count(app_species)
app_ab = data.frame("App", app_ab)
names(app_ab)<-c("ListID", "SpeciesID", "Freq")
app_ab

article_ab = count(article_species)
article_ab = data.frame("Plants_article", article_ab)
names(article_ab)<-c("ListID", "SpeciesID", "Freq")
article_ab
```

Matrifying the csv files:

```{r}
all_ab = rbind(welikia_ab, app_ab, article_ab)
all_matrix = matrify(all_ab)
all_matrix
```

Diversity/similarity indices:

```{r}
simpson<-diversity(all_matrix, index = "simpson")
shannon<-diversity(all_matrix, index = "shannon")
print("Simpson Index: ", quote = FALSE)
simpson
print("Shannon Index: ", quote = FALSE)
shannon
jaccard = vegdist(all_matrix, "jaccard")
print("Jaccard Index: ", quote = FALSE)
jaccard
bray = vegdist(all_matrix, "bray")
print("Bray Index: ", quote = FALSE)
bray
```

The three plant lists all have similarly high diversity indices. The Jacard and Bray indices are both "1" which means that there are no shared species between the lists.
